---
title: "Live Coding Quiz 2 Master"
author: "SY"
date: "10 March 2022"
output:
  word_document: default
  html_document:
    theme: readable
---

# Overview

Question 1: Import data and show first six lines. Code is already written. This tests a student's data management practices/understanding of troubleshooting.

Question 2: Summary statistics for central tendency and variation

Question 3: Troubleshoot an error for a data summary table

Question 4: Complete the steps for a graph

Question 5: Render to word doc (DOCX)

Three versions (A, B, C):

  + A. 11:35 and 11:50 time slots
  + B. 12:05 and 12:20 time slots
  + C. 12:35 time slot
  
Some options for alternatives that are correct are marked as comments; not exhaustive.

***

# Task 1:  Import data set and show first 6 lines of data frame

(1 point)

```{r echo = FALSE, results = 'hide'}
cheetah <- read.csv("Cub_probability_of_independence.csv", stringsAsFactors = TRUE, fileEncoding = "UTF-8-BOM")
head(cheetah)
```

# A Task 2:  What is the mode of Litter_fate?

(1 point)

```{r}
# option 1
tab <- table(cheetah$Litter_fate) 
tab # not required
names(tab)[which(tab==max(tab))]

# option 2
mode = function(x){ 
    ta = table(x)
    tam = max(ta)
    if (all(ta == tam))
         mod = NA
    else
         if(is.numeric(x))
    mod = as.numeric(names(ta)[ta == tam])
    else
         mod = names(ta)[ta == tam]
    return(mod)
}

mode(cheetah$Litter_fate)
```

# A Task 3:  Troubleshoot this error. 

(1 point)

The goal is to get the prey density median values of the different Litter_fate possibilities

```{r}
#preytable <- tapply(cheetah$Litter_fate, cheetah$Prey, median)
preytable <- tapply(cheetah$Prey, cheetah$Litter_fate, median)
preytable
```

# A Task 4: Complete the graph

(1 point)

The goal is to have a barplot of the median prey density (Prey, individuals/km2) on the y-axis and Litter_fate on the x-axis. Make sure the graph has all of the proper elements of a scientific graph.

```{r}
bar <- barplot(preytable, ylab = "Median Prey Density", xlab = "Litter Fate", main = "How Prey Denisty affects Litter Fate", ylim = c(0, 50), col = c("pink","yellow", "orange")) # needs labels, title; optional:  adjusting ylim and colors
```
***

# B Task 2: What is the range of Lion?

(1 point)

```{r}
# option 1
max(cheetah$Lion) - min(cheetah$Lion)

# option 2
diff(range(cheetah$Lion))
```

# B Task 3: Troubleshoot this error. 

(1 point)

The goal is to find out the total number of cheetah cubs that survived in each litter (40 litters)

```{r}
# Ltable <- table(cheetah$Survive[cheetah$Litter_ID])
Ltable <- tapply(cheetah$Survive, cheetah$Litter_ID, sum) 

Ltable
```

# B Task 4: Complete the graph

(1 point)

The goal is to have an XY scatterplot of the prey densities (individuals per km2) on the y-axis and density of lions (individuals per km2) on the x-axis, with the those that survived in transparent blue and those who did not survive in transparent black. Make sure the graph has all of the proper elements of a scientific graph.

```{r}
plot(cheetah$Lion[cheetah$Survive==1],
     cheetah$Prey[cheetah$Survive==1],
     pch = 19,
     col = adjustcolor("blue", alpha = 0.5),
     xlab= "Lion Density",
     ylab= "Prey Density",
     main= ""
     ) # needs labels, title
points(cheetah$Lion[cheetah$Survive==0],cheetah$Prey[cheetah$Survive==0], pch = 19, col = adjustcolor("black", alpha = 0.5))
legend("bottomleft", c("survived", "did not survive"), 
       pch = 19, 
       col = c("blue", "black"), 
       bty = "n")
```
***

# C Task 2: What is the skewness of Cheetah (individuals per km2)? Interpret.

(1 point)

```{r}
#install.packages("e1071")
library(e1071)
skewness(cheetah$Cheetah)
```

> Positive value, right skew

# C Task 3: Troubleshoot this error. 

(1 point)

The goal is to make a summary table of the mean Cheetah (individuals per km2) for each Litter_fate

```{r}
#CHtable <- aggregate(cheetah$Litter_fate, cheetah$Cheetah, mean)
CHtable <- aggregate(cheetah$Cheetah, list(cheetah$Litter_fate), mean)
names(CHtable) <- c("Litter_fate", "Cheetah_density")
CHtable
```

# C Task 4: Complete the graph

(1 point)

The goal is to have an XY scatterplot of Cheetah (cheetah density; individuals per km2) on the y-axis and Lion (lion density; individuals per km2) on the x-axis. Make sure the graph has all of the proper elements of a scientific graph.

```{r}
plot(cheetah$Lion, cheetah$Cheetah, 
     pch = 19,
     xlab= "Lion Density (individuals/km2)",
     ylab= "Cheetah Density (individuals/km2",
     main= "Effect of Predator Density on Prey Density"
     ) # needs labels, title

```


***

# Task 5

(1 point)

Knit this RMD to DOCX, and upload (1) Revised RMD and (2) DOCX to Canvas in the Quiz 2 assignment in the next five minutes.

If you are unable to render your Rmd file into a DOCX because of a problem in a particular code chunk, you may use {r eval = FALSE} instead of {r} for that code chunk. This should skip the chunks with the errors.