---
title: "Live Coding Quiz 1 Master"
author: "SY"
date: '2/15/2023 '
output:
  word_document: default
  html_document:
    df_print: paged
---

# Overview of Quiz 1

Three categories of coding tasks:  

1. information about data frame/objects
2. extract a specific element/subset (including logical operators)
3. calculation/mathematical operator

Three versions:

  + A. 11:00 and 11:13 time slots
  + B. 11:26 and 11:39 time slots
  + C. 11:52 and 12:05 time slots

# Import data set and load packages

```{r}
#install.packages("tidyverse")
library(tidyverse)

dw <- read.csv("dogORwolf.csv", header = T, stringsAsFactors = TRUE, fileEncoding = "UTF-8-BOM")
head(dw)
```

## A

```{r}
dim(dw)
oe2 <- dw$olfactory_exploration[dw$age_yrs > 2]
mean(oe2, na.rm = T) #na.rm is optional

```

## B 

```{r}
CWD <- subset(dw, breed == "CWD") 
# OR
CWD <- dw %>% 
  filter(breed == "CWD")

dim(CWD)

dw$total_exploration <- dw$visual_exploration + dw$olfactory_exploration
# OR
dw <- dw %>% 
  mutate(total_exploration = visual_exploration + olfactory_exploration)

```

## C

```{r}
is.numeric(dw$manipulation)
is.factor(dw$manipulation)

fdogs <- subset(dw, sex == "f")
# OR
fdogs <- dw %>% 
  filter(sex == "f")


max(dw$look_elsewhere)
# OR
summary(dw$look_elsewhere)
```

***

# Version A

## A Task 1:  How many rows and columns are in the dataframe?

(1 point)

```{r}
dim(dw)
```

## A Task 2:  Use square brackets to make a vector (array) called oe2 and put into it all of the values for olfactory_exploration for dogs older than 2 years (age_yrs)

(1 point)

```{r}
oe2 <- dw$olfactory_exploration[dw$age_yrs > 2]
```

## A Task 3: Obtain the mean of oe2

(1 point)

```{r}
mean(oe2, na.rm = T) #na.rm is optional
```

***

# Version B

## B Task 1: Make a subset (all columns) of the *dw* data frame called CWD, containing the breed CWD

(1 point)

```{r}
CWD <- subset(dw, breed == "CWD") 
# OR
CWD <- dw %>% 
  filter(breed == "CWD")
```

## B Task 2: How many rows and columns are in the *CWD* dataframe that you just made?

(1 point)

```{r}
dim(CWD)
```

## B Task 3: Make a new column in the *dw* dataframe called total_exploration and put the sum of visual_exploration and olfactory_exploration in it

(1 point)

```{r}
dw$total_exploration <- dw$visual_exploration + dw$olfactory_exploration
# OR if they are comfortable with Getting to Know the Data
dw <- dw %>% 
  mutate(total_exploration = visual_exploration + olfactory_exploration)
```

***

# Version C

## C Task 1: What kind of data is in the manipulation variable?

(1 point)

```{r}
is.numeric(dw$manipulation)
is.factor(dw$manipulation)
```

## C Task 2:  Make a subset (all columns) of the *dw* data frame called fdogs, containing only female dogs (sex is f)

(1 point)

```{r}
fdogs <- subset(dw, sex == "f")
# OR
fdogs <- dw %>% 
  filter(sex == "f")
```

## C Task 3:  What is the largest value of look_elsewhere for the *dw* data frame?

(1 point)

```{r}
max(dw$look_elsewhere)
# OR
summary(dw$look_elsewhere)
```

***

# Task 4

(1 point)

How do you feel about this quiz? On the line below, type your response using the comment formatting so it doesn't show up on the rendered Word document.


# Task 5

(1 point)

Knit this RMD to DOCX, and upload (1) Revised RMD and (2) DOCX to Canvas in the Quiz 1 assignment in the next five minutes.

If you are unable to render your Rmd file into a DOCX because of a problem in a particular code chunk, you may use {r eval = FALSE} instead of {r} for that code chunk. This should skip the chunks with the errors.
